The Community Land Model, version 5.1 \\ One-at-a-time Parameter Perturbation Ensemble

### 2023-12-19 ms/ms.tex

\subsection{Experimental Design}
Identification of the complete parameter set of a land surface model is in itself a non-trivial exercise, as in practice, many empirical constants were hard-coded and not amenable to systematic perturbation. For the purposes of this activity, we identified a broad set of 193 CLM-BGC parameters, and enabled their modification via the model input files. Several domains were not perturbed, including crops, CH$_4$, VOC, and dust emissions, urban properties, and river flow parameters. The crop model was not utilized as our initial focus is on unmanaged land.

@@

\label{sect:exps}
 \begin{table}[h]
 \caption{Forcing Scenarios}
 \centering
 \begin{tabular}{l c c c c}
 \hline
  Name  & Meteorology & CO$_2$ (ppmv) & N addition & Description \\
 \hline
   CTL2010  & 2005-2014 & 367 & - & control experiment\\
   C285        & 2005-2014 & 285 & - & low CO$_2$ \\
   C867        & 2005-2014 & 867 & - & high CO$_2$ \\
   AF1855    & 1851-1860 & 367 & - & pre-industrial climate \\
   AF2095    & 2091-2100 & 367 & - & late century climate (SSP3-7.0) \\
   NDEP      & 2005-2014 & 367 & 5g N m$^-2$  y$^-1$ & enhanced nitrogen deposition \\
 \hline
 \end{tabular}
 \label{tab:exps}
 \end{table}

@@

A total of 193 parameters were targeted for sensitivity testing across a wide variety of CLM subcomponents (Figure \ref{fig:params}).

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/main/bar.png}
\caption{193 parameters were perturbed across the various domains of the land model.}
\label{fig:params}
\end{figure}

Given the challenging size of the CLM5-BGC parameter space, and the need to probe and document parametric uncertainties within operational land surface models, a major goal of this project was to develop a fast configuration of CLM5-BGC that would allow for a large number of simulations given the available computational resources.
Combining the SASU spinup approach with our sparsegrid formulation (see Section \ref{sect:sg} for details) yielded a configuration approximately 500 times cheaper than the 1-degree configuration most often used for CLM simulations (Figure \ref{fig:sims}).
There are 22648, 5666, and 1764 land grid cells in standard 1$^{\circ}$, 2$^{\circ}$, and 4$^{\circ}$x5$^{\circ}$ CLM simulations, respectively, as compared to just 400 grid cells in the sparsegrid.
Likewise, whereas previous spin-up methodologies required 1500 years or longer to satisfy equilibrium criteria, the SASU approach enabled by matrix representation yielded satisfactory spin-up efficiency for our experiment within 140 years.

@@

A small fraction of parameters tend to explain a large amount of the ensemble variance (Figure \ref{fig:variance}) for any given variable or metric. For example, the same four parameters explain up to 90\% of the canopy evaporation variance across all the forcing scenarios (Supp Fig \ref{supp:fcev}). In general,most of the parameter perturbations had a relatively small impact on any given output variable. As such, the effective parameter space dimensionality for any given CLM process is likely much smaller than would be expected from having 193 independent parameters, reflecting the fact that land models like CLM comprise representations of numerous interconnected domains. While the principle of land models is that these are all interlinked via complex feedback processes, many parameters nonetheless have impacts that are mostly limited to their own domain.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/variance.pdf}
\caption{Cumulative fraction of variance explained by the most influential parameters on gross primary production and canopy evaporation. For example, approximately 96\% of canopy evaporation variance is explained by the first ten parameters regardless of forcing, whereas between 74 and 92\% of GPP variance is explained by the first ten parameters. Only the top 50 parameters are shown (of 193) to improve the visualization. Note that parameter rankings differ by variable.}
\label{fig:variance}
\end{figure}

@@

In this project, we identified and perturbed 193 CLM parameters to create a large one-at-a-time PPE. This ensemble is useful for understanding parametric controls on CLM processes. The software and analysis tools generated to enable and utilize this experiment will greatly reduce the burden of generating future PPEs and parameter sensitivity experiments.

There were several barriers to perturbing the full set of CLM5-BGC parameters. First, many parameters had not been officially identified as such. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. Although in this experiment we perturbed a large number of parameters (193) across a variety of CLM processes (Figure \ref{fig:params}), this still does not cover the full set of CLM parameters, as some processes were not included, such as crop phenology and management. In carrying out this process of parameter identification, we  likewise unearthed many nuances in  the epistemology of what  `parameters' are in the context of Earth system models. Defining the parameters within a given model structure can be somewhat subjective, such that it would be difficult to collate a comprehensive or definitive set of parameters for a model like CLM.

@@

For this experiment we opted for a one-at-a-time perturbation strategy, testing a minimum and maximum value for each parameter. We found that parameter effects could be quite large, in some cases exceeding the effects of our various forcing scenarios (Figure \ref{fig:ranges}). That said, the majority of parameter perturbations had small effects for any specific model variable or metric, such that a majority of simulations were clustered around the default simulation. In the case of canopy evaporation, for example, more than 95\% of the ensemble variance could be explained by the ten most influential parameters (Figure \ref{fig:variance}). This indicates that there may be tractable parameter estimation sub-problems if global calibration of 193 parameters proves unattainable.

>>>>

\subsection{Experimental Design}
Identification of the complete parameter set of a land surface model is in itself a non-trivial exercise, as in practice, many empirical constants were hard-coded and not amenable to systematic perturbation. For the purposes of this activity, we identified a broad set of 211 CLM-BGC parameters, and enabled their modification via the model input files. Several domains were not perturbed, including crops, CH$_4$, VOC, and dust emissions, urban properties, and river flow parameters. The crop model was not utilized as our initial focus is on unmanaged land.

@@

\label{sect:exps}
 \begin{table}[h]
 \caption{Forcing Scenarios}
 \centering
 \begin{tabular}{l c c c c}
 \hline
  Name  & Meteorology & CO$_2$ (ppmv) & N addition & Description \\
 \hline
   CTL2010  & 2005-2014 & 367 & - & control experiment\\
   C285        & 2005-2014 & 285 & - & low CO$_2$ \\
   C867        & 2005-2014 & 867 & - & high CO$_2$ \\
   AF1855    & 1851-1860 & 367 & - & pre-industrial climate \\
   AF2095    & 2091-2100 & 367 & - & late century climate (SSP3-7.0) \\
   NDEP      & 2005-2014 & 367 & 5g N m$^{-2}$  y$^{-1}$ & enhanced nitrogen deposition \\
 \hline
 \end{tabular}
 \label{tab:exps}
 \end{table}

@@

Given the challenging size of the CLM5-BGC parameter space (Figure \ref{fig:params}), and the need to probe and document parametric uncertainties within operational land surface models, a major goal of this project was to develop a fast configuration of CLM5-BGC that would allow for a large number of simulations given the available computational resources.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/main/bar.png}
\caption{211 parameters were perturbed across the various domains of the land model.}
\label{fig:params}
\end{figure}

Combining the SASU spinup approach with our sparsegrid formulation (see Section \ref{sect:sg} for details) yielded a configuration approximately 500 times cheaper than the 1-degree configuration most often used for CLM simulations (Figure \ref{fig:sims}).
There are 22648, 5666, and 1764 land grid cells in standard 1$^{\circ}$, 2$^{\circ}$, and 4$^{\circ}$x5$^{\circ}$ CLM simulations, respectively, as compared to just 400 grid cells in the sparsegrid.
Likewise, whereas previous spin-up methodologies required 1500 years or longer to satisfy equilibrium criteria, the SASU approach enabled by matrix representation yielded satisfactory spin-up efficiency for our experiment within 140 years.

@@

A small fraction of parameters tend to explain a large amount of the ensemble variance (Figure \ref{fig:variance}) for any given variable or metric. For example, the same four parameters explain up to 90\% of the canopy evaporation variance across all the forcing scenarios (Supp Fig \ref{supp:fcev}). In general,most of the parameter perturbations had a relatively small impact on any given output variable. As such, the effective parameter space dimensionality for any given CLM process is likely much smaller than would be expected from having 211 independent parameters, reflecting the fact that land models like CLM comprise representations of numerous interconnected domains. While the principle of land models is that these are all interlinked via complex feedback processes, many parameters nonetheless have impacts that are mostly limited to their own domain.

@@

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/variance.pdf}
\caption{Cumulative fraction of variance explained by the most influential parameters on gross primary production and canopy evaporation. For example, approximately 96\% of canopy evaporation variance is explained by the first ten parameters regardless of forcing, whereas between 74 and 92\% of GPP variance is explained by the first ten parameters. Only the top 50 parameters are shown (of 211) to improve the visualization. Note that parameter rankings differ by variable.}
\label{fig:variance}
\end{figure}

@@

In this project, we identified and perturbed 211 CLM parameters to create a large one-at-a-time PPE. This ensemble is useful for understanding parametric controls on CLM processes. The software and analysis tools generated to enable and utilize this experiment will greatly reduce the burden of generating future PPEs and parameter sensitivity experiments.

There were several barriers to perturbing the full set of CLM5-BGC parameters. First, many parameters had not been officially identified as such. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. Although in this experiment we perturbed a large number of parameters (211) across a variety of CLM processes (Figure \ref{fig:params}), this still does not cover the full set of CLM parameters, as some processes were not included, such as crop phenology and management. In carrying out this process of parameter identification, we  likewise unearthed many nuances in  the epistemology of what  `parameters' are in the context of Earth system models. Defining the parameters within a given model structure can be somewhat subjective, such that it would be difficult to collate a comprehensive or definitive set of parameters for a model like CLM.

@@

For this experiment we opted for a one-at-a-time perturbation strategy, testing a minimum and maximum value for each parameter. We found that parameter effects could be quite large, in some cases exceeding the effects of our various forcing scenarios (Figure \ref{fig:ranges}). That said, the majority of parameter perturbations had small effects for any specific model variable or metric, such that a majority of simulations were clustered around the default simulation. In the case of canopy evaporation, for example, more than 95\% of the ensemble variance could be explained by the ten most influential parameters (Figure \ref{fig:variance}). This indicates that there may be tractable parameter estimation sub-problems if global calibration of 211 parameters proves unattainable.
